---
title: "Statistical Learning with High-Dimensional Data"
author: "Ariel Nataf"
subtitle: DSTI A20 Cohort
output:
  html_document:
    df_print: paged
---

For some unknown reason to me the plot function doesn't show me the clusters colors.
I have to knit the document to see the colored result it took me a lot of time to figure this way to avoid the problem

#Exercise 3

## 3,1 Loading the Data
```{r}
load("/Users/arielnataf/Desktop/DSTI/SLHD/Velib.Rdata")
```
## 3.2 Pretreatment et descriptive analysis

We look at the documentation first
```{r}
?velibCount
```

We find very informative metadata:
> The format is:
> - data: the nb of available bikes of the 1189 stations at 181 time points.
> - position: the longitude and latitude of the 1189 bike stations.
> - dates: the download dates.
> - bonus: indicates if the station is on a hill (bonus = 1).
> - names: the names of the stations.

We also look for any missing data
```{r}
is.na(Velib)
```

There isn't any.

```{r}
# We keep only the mean number of velib at each station
Velib_mean <- rowMeans(Velib$data)
data = cbind(Velib$position, Velib_mean)
data = cbind(Velib$bonus, data)

data
```


# 3.3 Data visualization
```{r}
X = data
?princomp
# The usual manner to do PCA in R
pca = princomp(X)
biplot(pca,col=c("red","blue"))
?biplot
```
Looking at the PCA, 
• we can see that longitude and latitude arrows stay in the middle little influence
• Velib$bonus (on a hill) goes to the right
• The mean number of available velibs goes to the left

Velib$bonus and velib at date are opposite. We can guess a hill has an impact on the usage of a velib


## 3.4 Clustering

### 3.4.1 Hierarchical clustering
```{r}
D = dist(data)
par(mfrow=c(2,2))
hc1 = hclust(D,method = "complete"); plot(hc1)
hc2 = hclust(D,method = "single"); plot(hc2)
hc3 = hclust(D,method = "centroid"); plot(hc3)
hc4 = hclust(D,method = "ward.D2"); plot(hc4)
```
Complete seems to be the best method looking at the dendograms

```{r, results='hide', message=FALSE}
hc1 = hclust(D,method = "complete")
cl1 = cutree(hc4,k = 2)
cl1
```

We looks at clusters with all variables
```{r}
plot(data,col=cl1,pch=19)
```
clusters depending of hill and the mean numbers are clearly there

Looking at the clusters on a pretty map:
```{r}
#install.packages("leaflet")
library(leaflet)

palette = colorFactor("RdYlBu", domain = NULL)
leaflet(X) %>% addTiles() %>%
    addCircleMarkers(radius = 3,
                     color = palette(cl1),
                     stroke = FALSE, fillOpacity = 0.9)
```
very pretty, I can see my home

On the map we can guess the second cluster is the hills.

### 3.4.2 k-means

```{r}
#install.packages("mclust")
library(mclust)
```

```{r}
out = Mclust(data[c("Velib_mean", "Velib$bonus")],G=1:10,modelNames = "EII")
plot(out, what = 'BIC')
```

We only use 2 clusters
```{r}
out2 = kmeans(data, centers = 2, nstart = 10)
plot(data,col=out2$cluster,pch=19)
```


```{r}
plot(data[,c(1,4)],col=out2$cluster,pch=19)
```

Comparison with kmeans and hclust on mean number of velib on a station and hills
```{r}
par(mfrow = c(1,2))
D = dist(data)
hc = hclust(D,method = "complete"); out.hc = cutree(hc,2)
plot(data[,c(1,4)],col=out.hc,pch=19,cex = 3,main='HC')
out.km = kmeans(data, centers = 2, nstart = 10)
plot(data[,c(1,4)],col=out.km$cluster,pch=19,cex = 3,main='kmeans')
```
it looks similar 


## 3.5 Summary
People don't use as much velibs on top of a hill than down.
Otherwise west/north/west/east don't impact very much.
